Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-18 inference.tex

@@

The correspondence between the inference rules and the RG judgment of
\S~\ref{sec:reasoning} is stated thus\footnote{The proof can be found in
the supplementary material.}:
\begin{theorem}
\label{thm:inference-sound}
  \emph{Forall} i,$R$,$I$,$c$,$\F$, if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\\vspace*{-0.2cm}
  \begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=\emptyset \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl = \F(\emptyset,\stg)}
  \end{array}
  \end{smathpar}
\end{theorem}
\noindent where the set expression $\F(\emptyset,\stg)$ has a natural
interpretation in the theory of sets.
% Where, ${\sf true} = \lambda(\stl,\stg).~true$ and $\stable(\R,I)
% \Leftrightarrow \forall\stl,\stg,\stg'. I(\stg) \conj
% \R(\stl,\stg,\stg') \Rightarrow \I(\stg')$.

>>>>

The correspondence between the inference rules and the RG judgment of
\S~\ref{sec:reasoning} is stated thus\footnote{The proof can be found in
the supplementary material.}:
\begin{theorem}
\label{thm:inference-sound}
  \emph{Forall} i,$R$,$I$,$c$,$\F$, if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\\vspace*{-0.2cm}
  \begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=\emptyset \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl = \F(\emptyset,\stg)}
  \end{array}
  \end{smathpar}
\end{theorem}
\noindent where the set expression $\F(\emptyset,\stg)$ has a natural
interpretation in the theory of sets.

\begin{proof}
  We only provide a brief proof sketch here. We prove a slightly
  stronger statement :

  \emph{Forall} i,$R$,$I$,$c$,$\F$,$s$ if $\stable(\R,I)$ and $c
  \elabsto \F$, then:\\

  \begin{smathpar}
    \begin{array}{c}
    \R \vdash \hoare{\lambda(\stl,\stg).~\stl=s \conj
    I(\stg)}{c}{\lambda(\stl,\stg).\stl = \F(s,\stg)}
    \end{array}
  \end{smathpar}

  We use structural induction on $c$ to prove the statement. Assume
  that $P \Leftrightarrow \lambda(\stl,\stg).~\stl=s \conj I(\stg)$
  and $Q \Leftrightarrow \lambda(\stl,\stg).\stl = \F(s,\stg)$. For
  each type of $c$, our strategy is to show the premises of the
  corresponding RG-rule. The base cases correspond to the INSERT,
  UPDATE and DELETE statements. In all these cases, the form of $F$
  directly corresponds to the condition linking the pre-condition $P$
  and post-condition $Q$ in the premises of the RG-rules. If
  $\stable(\R, F)$, then $\stabilize{F} = F$ and hence the proof
  directly follows. Otherwise, $\stabilize{F} = \lambda (\stl,
  \stg).\existsl(\stg', I(\stg'), F(\stl, \stg'))$. Since the
  pre-condition $P(\stl, \stg)$ asserts $I(\stg)$, any state which
  obeys $P$ provides the existential $\stg'$ in the above set
  expression, and hence $F$ applied on that state matches the
  condition linking $P$ and $Q$. The stability of $P$ with respect to
  $\R$ follows directly from the stability of $I$.

  The proofs for SELECT, SEQ and IF-THE-ELSE directly use the inductive
  hypothesis to infer the RG-judgements present in the premises of
  their corresponding RG-rules. The interesting case is the FOREACH
  statement, for which we use the loop invariant $\psi(\stl, \stg)
  \Leftrightarrow \stl = s \cup y \bind (\lambda z.~\F(s, \stg))$,
  (where assuming that $c$ is the body of the loop, $c \elabsto F$) to
  prove all the premises of \rulelabel{RG-Foreach}. Using the same
  notation as the rule $\rulelabel{RG-Foreach}$, $y$ refers to the
  records already processed in previous iterations of the loop, while
  $z$ refers to the record being processed in the current iteration.
  At the beginning of the loop $[\phi/y]\psi(\stl, \stg)$ just reduces
  to $\stl = s$ which is implied by the pre-condition $P$. From the
  inductive hypothesis, we can infer that each iteration corresponds
  to the application of $F$. Since all iterations are assumed to be
  independent of each other, and $z$ is bound to a record in $x$ for
  each iteration, we can infer that at the end of every iteration the
  loop invariant $[y \cup \{z\}/y]\psi$ will be satisfied. 
\end{proof}

@@
